### 2009-05-21 doc/tech_doc/infosys/infosys_technical.tex
Authorization configuration with examples. Use architecture figure from Ivan. Better name of document.

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {},
  pdftitle = {Documentation of the ARC Information System},
  pdfsubject = {},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

\begin{figure}[ht]
\centering{{\scalebox{0.6}{\includegraphics{InfosysArchitecture.png}}}
\caption{\label{fig:infosys_architecture}Architecture of Information System} }
\end{figure}

Information System of ARC middleware is composed of 3 main parts. The Information System Indexing Services (ISIS) for a set of information containers. Every generic ARC Service pushes information about itself into nearby ISIS container/service hence registering itself to Information System. The information stored in every ISIS is propagated among all ISIS services. Clients can query any nearby ISIS service for registered information in order to perform discovery of services which posses specific properties.

@@

\begin{table}{h}
\caption{ARC Policy Actions of ISIS}
\label{tab:policy_actions}
\begin{tabular}{|c|c|l|}
\hline
\textbf{Group}&\textbf{Action}&\textbf{AttributeId}\\
\hline
Request from ISIS&\textit{isis}&http://www.nordugrid.org/schemas/policy-arc/types/isis/operation\\
\hline
Request from Service&\textit{service}&http://www.nordugrid.org/schemas/policy-arc/types/isis/operation\\
\hline
Request from generic client&\textit{client}&http://www.nordugrid.org/schemas/policy-arc/types/isis/operation\\
\hline
\end{tabular}
\end{table}

@@

@@

Following elements are defined inside configuration element if the Registrant:
\begin{description}
\item{URL}            <!-- URL specifies contact endpoint of a bootstrap Information
                 Registration service. Further ISIS addresses will be queried
                 from this service. -->
\item{KeyPath, CertificatePath, ProxyPath and CACertificatesDir}
            <!-- Optional KeyPath, CertificatePath, ProxyPath and CACertificatesDir are
                 paths to files storing X509 credentials used for establishing connections. -->
\item{Retry}            <!-- Retry count. Specifies how many times have an ISIS to timeout to handle it as
                 unavailable. If there is no value specified the default value: 5 will be used. -->
            <xsd:element name="Retry" type="xsd:string" minOccurs="0" maxOccurs="1"/>

>>>>

\documentclass{book}
\usepackage{graphicx}                              %for PNG images (pdflatex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%\usepackage{trackchanges}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2CM}
\hypersetup{
  pdfauthor = {},
  pdftitle = {Documentation of the Web Service based ARC Information System},
  pdfsubject = {},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\hyphenation{preserve-Original}
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

\begin{figure}[ht]
 \centering{
  \scalebox{0.6}{\includegraphics{SystemOverview.png}}
  \caption{\label{fig:infosys_architecture}Architecture of Information System}
 }
\end{figure}

Information System of ARC middleware is composed of 3 main parts. The Information System Indexing Services (ISIS) form a set of information containers. Every generic ARC Service pushes information about itself into nearby ISIS container/service hence registering itself to Information System. The information stored in every ISIS is propagated among all ISIS services. Clients can query any nearby ISIS service for registered information in order to perform discovery of services which posses specific properties. Clients can also query ARC Services to obtain more detailed information about Service properties.

@@

\begin{table}[h]
\caption{ARC Policy Actions of ISIS}
\label{tab:policy_actions}
\begin{tabular}{|c|c|l|}
\hline
\textbf{Group}&\textbf{Action}&\textbf{AttributeId}\\
\hline
Request from ISIS&\textit{isis}&http://www.nordugrid.org/schemas/policy-arc/types/isis/operation\\
\hline
Request from Service&\textit{service}&http://www.nordugrid.org/schemas/policy-arc/types/isis/operation\\
\hline
Request from generic client&\textit{client}&http://www.nordugrid.org/schemas/policy-arc/types/isis/operation\\
\hline
\end{tabular}
\end{table}

@@

\subsection{Configuration}
\label{sub:isis_authorization_configuration}

\textit{For information about sophisticated authorization policies and how to deploy various Policy Decision Point entities please see "Security Framework of ARC1" document"~\cite{arcsec}.}

To restrict set of clients allowed to perform operations in ISIS service proper authorization policy is needed. Let's assume ISIS is operating over TLS connection and all participants posses X.509 certificates with following subject names:
\begin{itemize}
\item /O=Grid/O=Test/CN=CLIENT - generic client entities
\item /O=Grid/O=Test/CN=SERVICE - generic service entities
\item /O=Grid/O=Test/CN=ISIS - all ISIS belonging to ISIS cloud
\end{itemize}

\textit{In NO way we suggest to use such setup. Real installation should use more sophisticated way to 
identify clients contacting ISIS service. For an infrastructure with quite static roles distribution for 
example we suggest to use Virtual Organization Management Service (VOMS) attributes embedded into X.509 
certificates representing participating entities.}

Below is an example policy made of 3 rules defined in lines 4-17, 18-29 and 30-39. Those define allowed 
behavior for ISIS, generic service and generic client. Lines 6-7, 20-21 and 32-33 specify attributes 
used to recognise type of connecting client. In this case those are subjects of X.509 certificates with 
values defined above. Lines 10-15, 24-27 and 36-37 specify allowed actions. One can see that this policy 
allows all operations to be performed by ISIS client. It limits operations allowed for generic service 
to "service" and "client" types. And client entities are allowed to perform "client" type operations only.

\begin{verbatim}
 1: <?xml version="1.0" encoding="UTF-8"?>
 2: <Policy xmlns="http://www.nordugrid.org/schemas/policy-arc" PolicyId="policy1"
 3:           CombiningAlg="Deny-Overrides">
 4:  <Rule RuleId="isis_to_isis" Effect="Permit">
 5:    <Subjects>
 6:      <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/identity"
 7:              >/O=Grid/O=Test/CN=ISIS</Subject>
 8:    </Subjects>
 9:    <Actions>
10:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
11:             >isis</Action>
12:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
13:             >service</Action>
14:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
15:             >client</Action>
16:    </Actions>
17:  </Rule>
18:  <Rule RuleId="service_to_isis" Effect="Permit">
19:    <Subjects>
20:      <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/identity"
21:              >/O=Grid/O=Test/CN=SERVICE</Subject>
22:    </Subjects>
23:    <Actions>
24:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
25:             >service</Action>
26:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
27:             >client</Action>
28:    </Actions>
29:  </Rule>
30:  <Rule RuleId="client_to_isis" Effect="Permit">
31:    <Subjects>
32:      <Subject AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/tls/identity"
33:              >/O=Grid/O=Test/CN=CLIENT</Subject>
34:    </Subjects>
35:    <Actions>
36:      <Action AttributeId="http://www.nordugrid.org/schemas/policy-arc/types/isis/operation"
37:             >client</Action>
38:    </Actions>
39:  </Rule>
40: </Policy>

This policy is not as restrictive as it could be in order to allow ISIS register themselves as ordinary 
services and to allow services which perform discovery of other services to behave like ordinary clients.

In order to activate policy it must be linked to ISIS service through Security Handle and Policy 
Decision Point emedded into configuration of ISIS. Below is minimal example configuration element of 
ISIS service. In lines 7-17 a set of Security Handler and Policy Decision Point is configured to handle
policies written in ARC Policy Language (described in ~\cite{arcsec}). Line 14 specifies that policy is 
read from file \textit{/opt/arc/etc/isis/policy.xml} every time new request is arrived. If policy is not 
satisfied then ISIS returns SOAP Fault instead of usual informative response.

\begin{verbatim}
 1: <Service 
 2:   xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
 3:   xmlns:isis="http://www.nordugrid.org/schemas/ArcConfig/2009/isis"
 4:   name="isis" id="isis">
 5:  <isis:endpoint>https://localhost:60000/isis</isis:endpoint>
 6:  <isis:DBPath>/opt/arc/share/isisdb</isis:DBPath>
 7:  <SecHandler
 8:    xmlsns:auth="http://www.nordugrid.org/schemas/SimpleListAuthZ"
 9:    name="arc.authz" id="authz" event="incoming">
10:   <auth:PDP
11:     xmlns:pdp="http://www.nordugrid.org/schemas/ArcPDP"
12:     name="arc.pdp">
13:    <pdp:PolicyStore>
14:     <pdp:Location type="file">/opt/arc/etc/isis/policy.xml</pdp:Location>
15:    </pdp:PolicyStore>
16:   </auth:PDP>
17:  </SecHandler>
18: </Service>

@@

Following elements are defined inside configuration element of the Registrant:
\begin{description}
\item{URL} specifies contact endpoint of a bootstrap ISIS. If neede further ISIS addresses will be 
queried from this service. This element is mandatory.
\item{KeyPath, CertificatePath, ProxyPath and CACertificatesDir} are paths to files storing X509 
credentials used for establishing connections. Those elements are optional and needed only if 
TLS communication is used.
\item{Retry} specifies how many times communication to ISIS have to fail/timeout to start treating is
as unavailable. The default value is 5.
\end{description}
